Infra-estrutura de Testes para Implementações de Referência do Standard ECMAScript

### 2022-05-27 paper.tex
draft introduction before meeting with the teacher

%
\begin{abstract}
The ECMAScript standard is the specification for JavaScript which is one of the most popular programming languages in the world. Implementing the standard is challenging in various ways, both due to its complexity and its scale. Test262 was created to alleviate the implementations difficulty by providing a conformance test suite that allows developers to test their implementations. There is the project MetaData262 v0 that calculates the metadata of the Test262 test suite. With this thesis, we will improve the MetaData262 v0 metadata calculations as well as build infrastructure to view the metadata through a web interface. The web interface will also allow to filter tests by some characteristics such as built-ins and version of the standard as well as display statistics of the metadata.

@@

@@

\begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
  1995 & JavaScript is launched and implemented in Netscape Navigator 2.0 beta 3.\\
  2005 & Thiemann presents the first try at defining a system of types possible for JavaScript ~\cite{Thiemann-2005}. The system detects the possible characteristics of an object and signals suspected type conversions. The work done does not cover important aspects of JavaScript, such as inheritance based on prototypes, and was not created along with an implementation at the time of the publication.\\
  2008 & Maffeis et al. proposes a first operational semantics for the language ECMAScript version 3. (ES3)~\cite{operational_semantics_javascript-2008}. This first semantics was used to analyze various properties of security of web applications.\\
  2010 & Guha et al.~\cite{essence_javascript-2010} defines a language $\lambda$JS, a formal calculation that encapsulates the fundamental characteristics of ES3. $\lambda$JS comes along an interpreter \emph{Racket}~\cite{Racket} and a translator from ES3 programs to expressions $\lambda$JS.\\
  2012 & Gardner et al.~\cite{Gardner-2012} adapts the logic ideas of separation to build a scalable Hoare logic for a subset of JavaScript, based of semantics faithful to ES3.\\
\end{tabular}
\vspace{1em}
\caption{Time line that contains the most relevant related work to the objectives of the project (1 of 2)}
\label{tab:related-work_timeline-1}
\end{table}
\vspace{-2em}

@@

  2012 & Politz et al.~\cite{Politz-2012} extends the language $\lambda$JS of ES3 to ES5 included, fir the first time, a formal semantics of JavaScript \texttt{property descriptors} and complete handling of the \texttt{eval} declaration.\\
  2014 & Bodin et al. presents \emph{JSCert}~\cite{JSCert-2014}, a formalization of the ES5 standard written in \emph{Coq}~\cite{Coq}, and \emph{JSRef} an executable interpreter of references extracted from \emph{Coq} to \emph{OCaml}~\cite{OCaml}, proving correct when related to \emph{JSCert} and tested using the test suite Test262~\cite{Test262}.\\
  2015 & Park et al. presents \emph{KJS}~\cite{KJS-2015}, a formal semantics for ES5, the most complete to that moment. The \emph{KJS}  was developed using the framework \emph{K}~\cite{Kframework}, a consolidated system of rewriting of terminology that supports various types of symbolic analyses based on \emph{reachability logic} (All-Path Reachability Logic \cite{All-Path_Reachability_Logic}). \emph{KJS} was tested carefully using the test suite, passing all the 2.782 tests of the language core. The \emph{KJS} can be executed symbolically, being able to be used for formal analyses and verification of ES5 programs.\\
  2015 & Gardner et al. \cite{JSCert_1Year-2015} extend the project \emph{JSCert} to support \emph{arrays} ES5. In order to do this, the authors connected the \emph{JSRef} to the implementation of the Array library of the V8 engine \cite{V8} from Google. Additionally, the authors evaluated the state of the project \emph{JSCert} at the time, giving a detailed analysis of the methodology as a whole and a detailed analysis of the tests that the project passed or failed.\\
  2017 & Fragoso Santos et al. creates the tool \emph{JaVert}~\cite{Javert-2017}: a tool for verifying JavaScript that allows semiautomatic reasoning about the fundamental properties of correctness of ES5 programs. JaVert does not assume any simplification of the language semantics, analyzing all the \emph{corner-cases} of the language. The load of necessary notation is substantial.\\
  2018 & Charguéraud et al. presents \emph{JSExplain}~\cite{JSExplain-2018}, an reference interpreter for the ES5 language that follows the specifications and produces an inspectable execution traces. It also includes an interface of code inspection that allows the programmer to execute his ES5 code step by step as well as the code of the ES5 interpreter, while executing ES5 programs.\\
  2020 & Sampaio et al. ~\cite{Javert-2020} extend the JaVerT tool to support automatic reasoning about JavaScript events. In order to do this, the authors developed a reference implementation of JavaScript promises directly in JavaScript. The authors used \colorbox{orange}{it} to test their reference implementation, passing 344 tests of the 474 total dedicated to JavaScript promises. Around 106 testes were excluded due to ES6 features that were not yet supported and to tests that require \emph{non-strict} mode.\\
  2020 & Jihyeok Park et al. presents JISET~\cite{JISET-2020} the first tool that synthesizes automatically the analyzer and that translates AST-IR directly from a specification of a given language.
\end{tabular}
\vspace{1em}
\caption{Time line that contains the most relevant related work to the objectives of the project (2 of 2)}
\label{tab:related-work_timeline-2}
\end{table}
% QUESTION it e' Test262??

>>>>

%
\begin{abstract}
The ECMAScript standard is the specification for JavaScript which is one of the most popular programming languages in the world. Implementing the standard is challenging in various ways, both due to its complexity and its scale. Test262 was created to alleviate the implementations difficulty by providing a conformance test suite that allows developers to test their implementations. Test262 has many tests and its hard to select the tests that belong to or need a feature in order to work, MetaData262 v0 was created to help with that. MetaData262 v0 calculates metadata of the tests in the Test262 test suite, making it available in MongoDB database. With this thesis, we will improve the MetaData262 v0 metadata calculations as well as build infrastructure to view the metadata through a web interface. The web interface will also allow to filter tests by some characteristics such as built-ins and version of the standard as well as display statistics of the metadata.

@@

% Para 2 --> implemetar JS e' dificil: complexidade tamanho
% Para 2.5 --> Test262 ajuda a implementar permitindo testar implementacoes --> test262 tambem e' grande dificil selecionar os testes apropriados
JavaScript is defined in the \texttt{ECMAScript} standard, a document written in English. The standard has be growing with each version, both in size and complexity. The growth of the standard is illustrated in Figure \ref{fig:paginas_standard_version}. This makes the JavaScript a language very hard to implement with many corner cases. \texttt{Test262} was created to aid the development of JavaScript engines, by providing a conformance test suite that developer can use to test their implementations. As the standard is extensive and complex and its conformance test suite needs to thoroughly test the standard implementation, then, the test suite it self becomes very extensive. The standard being complex and extensive also leads to many of its implementations being only partial.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.7\textwidth]{images/standard_pages_evolution.png}
    \caption{Evolution of ECMAScript standard's number of pages}
    \label{fig:paginas_standard_version}
\end{figure}

% Para 4 --> MetaData v0 calcula metadata --> disponibiliza mongoDB

When implementing a JavaScript engine it is very important to know the tests that the engine should pass for the feature that is being implemented. The selection of applicable tests for partial implementations is particularly difficult since tests for one feature may or may not use other features that are not implemented. \texttt{MetaData262 v0} appears to solve this issue, it processes the \texttt{Test262} suite and calculates metadata about the tests. \texttt{MetaData262 v0} has some particularly interesting properties such as a version of the standard that the test belongs to, the syntactic constructs that are used in the test, and the built-ins that are used by the test. The metadata of \texttt{MetaData262 v0} is made available in the form of a MongoDB database.

% --> eu vou melhorar metadata
% Para 5 --> Visualization system permite interagir com a metadata via web com interface simples --> filtrar; verdados varias maneiras; estatisticas
In this thesis we will improve \texttt{MetaData262 v0} in its calculations of exiting metadata as well as adding the new metadata related with the harness functions used. We will also be building web application that interacts with the MongoDB database, allowing users to filter tests by their fields displaying the metadata of the tests along side statistics of the metadata. The aim of the web application is to make the metadata easily accessible.

% Para 6 --> resumo do resto do relatorio
This document is structured in the following way. Section \ref{sec:Goals} has the objectives of this thesis. Section \ref{sec:Background} describes the field in which this thesis is inserted, describing the \texttt{ECMAScript} standard, followed by a description the \texttt{Test262} test suite, and a description of \texttt{MetaData262 v0}. Next is section \ref{sec:Related Work} that presents the scientific work done that relates to this thesis theme. In section \ref{sec:Design and Methodology} is the design and methodology that this thesis is implemented with. As for section \ref{sec:Evaluation and Planning}, it describes how we plan on evaluating the results of this thesis as well as its timeline. Finally, section \ref{sec:Conclusion} is where we present the main conclusions and a summary of this proposal.

@@

\begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
  1995 & JavaScript is launched and implemented in Netscape Navigator 2.0 beta 3.\\
  2005 & Thiemann presents the first try at defining a system of types possible for JavaScript ~\cite{Thiemann-2005}. The system detects the possible characteristics of an object and signals suspected type conversions. The work done does not cover important aspects of JavaScript, such as inheritance based on prototypes, and was not created along with an implementation at the time of the publication.\\
  2008 & Maffeis et al. proposes a first operational semantics for the language ECMAScript version 3. (ES3)~\cite{operational_semantics_javascript-2008}. This first semantics was used to analyze various properties of security of web applications.\\
  2010 & Guha et al.~\cite{essence_javascript-2010} defines a language $\lambda$JS, a formal calculation that encapsulates the fundamental characteristics of ES3. $\lambda$JS comes along an interpreter \emph{Racket}~\cite{Racket} and a translator from ES3 programs to expressions $\lambda$JS.\\
  2012 & Gardner et al.~\cite{Gardner-2012} adapts the logic ideas of separation to build a scalable Hoare logic for a subset of JavaScript, based of semantics faithful to ES3.\\
  2012 & Politz et al.~\cite{Politz-2012} extends the language $\lambda$JS of ES3 to ES5 included, fir the first time, a formal semantics of JavaScript \texttt{property descriptors} and complete handling of the \texttt{eval} declaration.\\
  2014 & Bodin et al. presents \emph{JSCert}~\cite{JSCert-2014}, a formalization of the ES5 standard written in \emph{Coq}~\cite{Coq}, and \emph{JSRef} an executable interpreter of references extracted from \emph{Coq} to \emph{OCaml}~\cite{OCaml}, proving correct when related to \emph{JSCert} and tested using the test suite Test262~\cite{Test262}.\\
  2015 & Park et al. presents \emph{KJS}~\cite{KJS-2015}, a formal semantics for ES5, the most complete to that moment. The \emph{KJS}  was developed using the framework \emph{K}~\cite{Kframework}, a consolidated system of rewriting of terminology that supports various types of symbolic analyses based on \emph{reachability logic} (All-Path Reachability Logic \cite{All-Path_Reachability_Logic}). \emph{KJS} was tested carefully using the test suite, passing all the 2.782 tests of the language core. The \emph{KJS} can be executed symbolically, being able to be used for formal analyses and verification of ES5 programs.\\
  2015 & Gardner et al. \cite{JSCert_1Year-2015} extend the project \emph{JSCert} to support \emph{arrays} ES5. In order to do this, the authors connected the \emph{JSRef} to the implementation of the Array library of the V8 engine \cite{V8} from Google. Additionally, the authors evaluated the state of the project \emph{JSCert} at the time, giving a detailed analysis of the methodology as a whole and a detailed analysis of the tests that the project passed or failed.
\end{tabular}
\vspace{2em}
\caption{Time line that contains the most relevant related work to the objectives of the project (1 of 2)}
\label{tab:related-work_timeline-1}
\end{table}

@@

  2017 & Fragoso Santos et al. creates the tool \emph{JaVert}~\cite{Javert-2017}: a tool for verifying JavaScript that allows semiautomatic reasoning about the fundamental properties of correctness of ES5 programs. JaVert does not assume any simplification of the language semantics, analyzing all the \emph{corner-cases} of the language. The load of necessary notation is substantial.\\
  2018 & Charguéraud et al. presents \emph{JSExplain}~\cite{JSExplain-2018}, an reference interpreter for the ES5 language that follows the specifications and produces an inspectable execution traces. It also includes an interface of code inspection that allows the programmer to execute his ES5 code step by step as well as the code of the ES5 interpreter, while executing ES5 programs.\\
  2020 & Sampaio et al. ~\cite{Javert-2020} extend the JaVerT tool to support automatic reasoning about JavaScript events. In order to do this, the authors developed a reference implementation of JavaScript promises directly in JavaScript. The authors used \colorbox{orange}{it} to test their reference implementation, passing 344 tests of the 474 total dedicated to JavaScript promises. Around 106 testes were excluded due to ES6 features that were not yet supported and to tests that require \emph{non-strict} mode.\\
  2020 & Jihyeok Park et al. presents JISET~\cite{JISET-2020} the first tool that synthesizes automatically the analyzer and that translates AST-IR directly from a specification of a given language.
\end{tabular}
\vspace{1em}
\caption{Time line that contains the most relevant related work to the objectives of the project (2 of 2)}
\label{tab:related-work_timeline-2}
\end{table}
% % QUESTION it e' Test262??
